---
title: "Survival Analysis: University Completion Rate"
output: html_notebook
---

Get the current Working directory 
```{r}
getwd()
```
Set a working directory 
```{r}
setwd("C:/Users/lotomej/Desktop/R/degree/survival")
```
read a csv
```{r}
data <- read.table("data.csv")
```

EXPLORATORY DATA ANALYSIS
```{r}
head(data)
```
```{r}
summary(data)
```

Converting courses, Gender and Reasons into categories
```{r}
data_copy <- data
data_copy$Course[data_copy$Course == 1] = "BMCS"
data_copy$Course[data_copy$Course == 2] = "BSSC"
data_copy$Course[data_copy$Course == 3] = "BTRE"
data_copy$Course[data_copy$Course == 4] = "BTAP"
data_copy$Course <- factor(data_copy$Course)

data_copy$Gender[data_copy$Gender == 1] = "M"
data_copy$Gender[data_copy$Gender == 2] = "F"
data_copy$Gender <-factor(data_copy$Gender)

data_copy$Reasons[data_copy$Reasons == 1] = "Academics"
data_copy$Reasons[data_copy$Reasons == 2] = "Health"
data_copy$Reasons[data_copy$Reasons == 3] = "Discipline"
data_copy$Reasons[data_copy$Reasons == 4] = "Finance"
data_copy$Reasons[data_copy$Reasons == 5] = "Graduated"
data_copy$Reasons <-factor(data_copy$Reasons)

sample(data_copy)
```

#Survial Analysis
#Insalling the packages needed
```{r}
install.packages(c("survival", "survminer"))
```
loading librarires
```{r}
library("survival")
library("survminer")
```
censoring 
```{r}
data$time <- data$Year*12
data$Status <- ifelse(data$"Year">4, 0, 1)
```

# Kaplan-Meir plots
KM plot for Gender
```{r}
fit <- survfit(Surv(time, Status) ~Gender, data=data)
print(fit)
```

```{r}
summary(fit)
```
```{r}
summary(fit)$table
```
# kaplan-Meier survival estimate
* n: total number of subjects in each curve.
* time: the time points on the curve.
* n.risk: the number of subjects at risk at time t
* n.event: the number of events that occurred at time t.
* n.censor: the number of censored subjects, who exit the risk set, without an event, at time t.
* lower,upper: lower and upper confidence limits for the curve, respectively.
```{r}
d <- data.frame(time = fit$time,
                n.risk = fit$n.risk,
                n.event = fit$n.event,
                n.censor = fit$n.censor,
                surv = fit$surv,
                upper = fit$upper,
                lower = fit$lower
)
head(d)
```
# Visualization of the survival curve for Gender
```{r}
ggsurvplot(fit, data=data,
           pval = TRUE, conf.int = TRUE,
           risk.table = TRUE,
           risk.table.col = "strata",
           linetype = "strata", 
           surv.median.line = "hv",
           ggtheme = theme_bw(),
           palette = c("#E7B800", "#2E9FDF"))

ggsurvplot(fit, data=data,                    
                pval = TRUE,            
                conf.int = TRUE,         
                conf.int.style = "step",  
                xlab = "Time in days",   
                break.time.by = 10,     
                ggtheme = theme_light(), 
                risk.table = "abs_pct",  
                risk.table.y.text.col = T,
                risk.table.y.text = FALSE,
                ncensor.plot = TRUE,      
                surv.median.line = "hv",  
                legend.labs = 
                  c("Male", "Female"),    
                palette = 
                  c("#E7B800", "#2E9FDF"))
```
KM plot for Course Group
```{r}
fit <- survfit(Surv(time, Status) ~Course, data=data)
print(fit)
```

```{r}
summary(fit)
```

```{r}
summary(fit)$table
```

KM Survival Estimate
```{r}
d <- data.frame(time = fit$time,
                n.risk = fit$n.risk,
                n.event = fit$n.event,
                n.censor = fit$n.censor,
                surv = fit$surv,
                upper = fit$upper,
                lower = fit$lower
)
head(d)
```

Visualization
```{r}
ggsurvplot(fit, data=data,
           pval = TRUE, conf.int = TRUE,
           risk.table = TRUE, 
           risk.table.col = "strata",
           linetype = "strata", 
           surv.median.line = "hv",
           ggtheme = theme_bw(), 
           palette = c("#E7B800", "#2E9FDF", "#b6d477", "#d477b9"))

ggsurvplot(fit, data=data,                    
           pval = TRUE,            
           conf.int = TRUE,         
           conf.int.style = "step",  
           xlab = "Time in days",   
           break.time.by = 10,     
           ggtheme = theme_light(), 
           risk.table = "abs_pct",  
           risk.table.y.text.col = T,
           risk.table.y.text = FALSE,
           ncensor.plot = TRUE,      
           surv.median.line = "hv",  
           legend.labs = c("BMCS", "BSSC", "BTRE", "BTAP"),    
           palette = c("#E7B800", "#2E9FDF", "#b6d477", "#d477b9"))
```

# Log rank Tests 
Test on Gender Groups
```{r}
surv_dif <- survdiff(Surv(time, Status) ~Gender, data=data)
surv_dif
```

Test on Course Groups
```{r}
surv_dif <- survdiff(Surv(time, Status) ~Course, data=data)
surv_dif
```

Test on the Reasons
```{r}
surv_dif <- survdiff(Surv(time, Status) ~Reasons, data=data)
surv_dif
```

# Complex Survival Curves
Plot survival curves by sex and facet by rx and adhere
```{r}
require("survival")
fit2 <- survfit( Surv(time, Status) ~ Gender + Course,
                 data = data )
ggsurv <- ggsurvplot(fit2, fun = "event", conf.int = TRUE,
                     ggtheme = theme_bw(), data=data)

ggsurv$plot +theme_bw() + 
  theme (legend.position = "right")+
  facet_grid(Gender ~ Course)
```

# COX MODEL 
Univariate Cox regression
Gender
```{r}
res_cox <- coxph(Surv(time, Status) ~ Gender, data = data)
res_cox
```
```{r}
summary(res_cox)
```

Course
```{r}
res_cox <- coxph(Surv(time, Status) ~ Course, data = data)
res_cox
```

```{r}
summary(res_cox)
```

Multivariate Cox regression analysis
```{r}
cox_ <- coxph(Surv(time, Status) ~ Age + Gender + Course, data =  data)
summary(cox_)
```

Plot the baseline survival function
```{r}
ggsurvplot(survfit(cox_), color = "#3F9FDF", data=data,
           ggtheme = theme_minimal())
```







































